The Origins of Congestion and\\Network Assisted End-to-End Congestion Control

### 2020-05-25 paper/paper.tex
Feedback

\begin{itemize}
\item 
\textit{Transmission delay} describes the delay of transmitting data packets between two internet nodes. Typically, packets sent from an origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth. Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be sent at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]

@@

Whereas processing delay is almost not noticeable and transmission delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce. There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used. Traffic in residential network nodes for example differs from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
\\Sizing buffers is a complex task. On the one hand read and write times should be as short as possible to minimize additional delay. On the other hand buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}. Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers. Additionally, a 2004 paper by Appenzeller et al. \cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance. Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
\\This isn't the only major cause of queuing delay. Without proper buffer management even nodes with well-sized buffers will experience performance loss. Proper queue management is able to compensate for the unpredictable fluidity of network traffic. It is a difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate. After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them from being lost \cite{gettys2012bufferbloat}. If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.

\subsection*{Bufferbloat}
Bufferbloat refers to frequently full and excessively large buffers in internet nodes. This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP \cite{gettys2012bufferbloat,chen2014bufferbloat}. Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss. In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node early in. With memory getting cheaper over the years, buffers saw an increase in size. This has the negative impact, that more packets can be queued before any are dropped, thus those adjustments of congestion avoidance protocols kick in later than usual. Therefore, buffers are being filled in the whole network which results in an extreme increase in network delay (\textit{latency}) and fluctuation (\textit{jitter}) \cite{gettys2012bufferbloat,staff2012bufferbloat,chen2014bufferbloat}.
\\So Services, which require low latency at a high data-rate, such as streaming, will experience a decrease in quality.

\section{Congestion Avoidance}
\label{sec:ca}
As already mentioned, congestion detection algorithms make up an important part in congestion avoidance protocols. \textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}. To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works. TCP lives on the transport layer. It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packages over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}. It is the interface between network components and computer programs. Each of those packets have different purposes, but underlay the same structure - the protocol header. The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]:
\begin{figure}
\centering
\begin{bytefield}[bitheight=2.2\baselineskip ,bitwidth=0.635\baselineskip]{32}
	\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
	\bitbox{32}{Sequence Number} \\
	\bitbox{32}{Acknowledgment Number} \\
	\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} &
\bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} &
\bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} &
\bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} &
\bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} &
\bitbox{16}{Window Size} \\
\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
\bitbox{32}{Options} \\
\bitbox{32}{Data}
\end{bytefield}
\caption{Basic structure of the TCP header. It is split up into multiple 32-bit blocks. The data block contains all the application data to be transmitted \cite{jacobson1992tcp,ietf-tcpm-rfc793bis-16,huston2000tcp}.}
\label{fig:TCP_Header}
\end{figure}
\begin{itemize}
\item \textit{Source Port} and \textit{Destination Port} specify the ports on which TCP should communicate with the application layer.

@@

\subsection{Perceiving Congestion}
\label{P_C_marker}
Both partners of a connection are both sender and receiver, as both transmit packets. In the following however, we will simplify this and specify one partner as the sender, the other as the receiver. The sender keeps track of the time it takes from when a package sent until the acknowledgment is received. This is also known as \textit{round-trip time} (RTT). Comparing the current RTT to previously measured RTTs can inform the sender about the current state of congestion. Some variances of TCP make use of this feature in their congestion avoidance algorithms, yet classic TCP does not \cite{huston2000tcp,jacobson1995congestion}. RTT also plays a very important role in the detection of packet loss, which further implies congestion. The sender side keeps track of the RTT while it is being measured and compares it to the \textit{retransmission timeout} (RTO) \cite{jacobson1992tcp}. RTO time is being calculated by estimating the mean and variance of the RTT. Whenever packets are not acknowledged within this RTO interval, the TCP sender regards this packet as lost and will retransmit its data. Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}. To implicitly notify TCP that congestion occurs, congested network nodes will occasionally drop packets after a certain threshold of queued packets has been exceeded. Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.

\subsection{Adjusting for Congestion}
\label{AfC}
To prevent congestion altogether, packet flow inside a network would have to be conservative, i.e. a new packet isn't being sent into the network before an old packet leaves the network. A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}. Whenever a connection is in equilibrium, congestion in the network cannot be caused by this TCP connection, thus also controlling congestion. TCPs congestion avoidance algorithm can be categorized into three sub-algorithms, that all have the goal to keep the connection in equilibrium, no matter how congestion was perceived. Here, a variable called \textit{congestion window} (cwnd) has an important role. It describes the maximum amount of packets sent during an average RTT. An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}. The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.

@@

\section{Network Assisted Congestion Control}
The problem that comes with classic TCP is that the congestion detection algorithms are very implicit. Packet loss for example may have different causes other than network congestion, such as simple cable interference, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay. Therefore, exactly determining whether the network is congested or not is crucial to keep a network stable. Here, the network layer comes into play. Because most communication should be hidden along the path of the packet, all network nodes will only communicate on the network layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
In 2001 a new method was standardized to explicitly notify TCP on congestion. \textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the amount of packets dropped. This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}. Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}]. Thus, congestion can be described explicitly resulting in better short- and long-term network performance. Finally, if a data packet has been marked, the according acknowledgment packet will also be marked to inform the sender of congestion \cite{ramakrishnan1999proposal,ramakrishnan2001addition,10.1145/205511.205512}.\\
To sum up so far, without ECN, classic TCP would just drive the network into congestion, and then recover from it. The network generally speaking will at least experience some jitter (fluctuating network delay). With ECN however, packets in congested nodes will be marked, explicitly informing the TCP sender that the network is congested. As a result, avoiding actions can be taken immediately. In addition, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion. Therefore, this central algorithm is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.

@@

\subsubsection*{XCP Sender}
As with TCP, the XCP sender maintains a congestion window, \textit{cwnd}, and an estimate of the round trip time, \textit{RTT}. On packet departure, the \textit{H\_cwnd} and \textit{H\_rtt} fields will be initialized with \textit{cwnd} and \textit{RTT} accordingly. The initialization value of the \textit{H\_feedback} field represents a desired send rate $r$:
\[H\_feedback \leftarrow \frac{r \cdot RTT - cwnd}{cwnd}\]
This is an improvement to TCPs adjustment algorithms as the desired rate can be reached after one RTT.\\
The other task an XCP sender has is to adjust the congestion window according to the networks feedback. When $s$ is the packet size, the congestion window variable would be adjusted by following formula:
\[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
In addition to responding to direct feedback, the XCP sender still needs to handle packet loss, though this is very similar to what TCP does \cite{katabi2002congestion}.

@@

\subsubsection*{XCP Router}
The main task of network nodes in XCP, or XCP routers, is to inform the XCP sender of the current state of congestion. This is done by computing feedback at every node. The objective of XCP is to prevent the queues building up to a point at which packets have to be dropped \cite{katabi2002congestion}.\\
Feedback is computed by two controllers. The \textit{Efficiency Controller} (EC) tries to maximize link utilization while minimizing queue lengths and packet drop rates. When the XCP routers link is underused, positive feedback should be sent. If the link is congested however, negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
To achieve efficiency and equilibrium, this feedback will be allocated to single packets as \textit{H\_feedback}. However, the EC does not decide which packets should carry the feedback. Hence, the \textit{Fairness Controller} (FC) comes into play. As TCP, FC relies on the AIMD principle [\nameref{AfC}]. If the EC calculated positive feedback, it will be split up between all packets equally. If the feedback is negative however, it will be divided between packets in proportion to their current send-rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
Decoupling the feedback computation allows XCP to quickly acquire the maximum send rate and achieve equilibrium, all while the network allocates bandwidth to all XCP senders equally \cite{katabi2002congestion}.

\subsection{Evaluation of Network Assisted Congestion Control}
As you can see, including the network layer to detect congestion can have great performance improvements. 'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}. Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency, and throughput \cite{katabi2002congestion}.
\\However, all this performance comes at a cost. Since network nodes have to actively detect congestion, ECN, XCP, and other network assisted congestion control protocols rely on active queue management in those nodes. Although there is some hardware being produced today that meet ECNs requirements, a large part of internet hardware isn't capable of doing so. Because any of those nodes can be congested, it would be counterintuitive to run those protocols, as congestion might or might not be detected, thus resulting in performance loss \cite{katabi2002congestion}. This is the reason you do not see those protocols in wide area networks (WAN) too often, let alone the internet.
\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements. Especially smaller networks with high traffic could definitely benefit from those technologies.

\subsection{DCTCP}
A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center. Optimizing those data centers networking capabilities can result in an increase in application productivity. In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
The general problem that appears in high-traffic environments is the so-called \textit{TCP Incast} problem. It describes a drastic reduction in application throughput. To understand where TCP incast originates, you fist need to understand how most data centers are built. A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}]. This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers. It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}. In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers. Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged. Besides, many data center switches have relatively small, shared buffers. Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind. To achieve this, DCTCP implements ECN. Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received. Therefore, excessive link underutilization can be prevented. In addition, the DCTCP receiver will only send one ACK packet for every $m$ consecutive data packet, where all $1,...,m$ packets before had the same CE-flag configuration. Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation. Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
\\Being able to react to congestion immediately enables DCTCP to reduce buffer overflows and timeouts, thus largely resolving incast problems. The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.

@@

>>>>

\begin{itemize}
\item 
\textit{Transmission delay} is the time it takes to send a packet from one node to another. More specifically, it describes the delay of transmitting data packets between two internet nodes depending on bandwidth and length of the connection. Typically, packets sent from an origin to a destination node are sent over various different internet nodes, all of which could have a different bandwidth. Therefore, a packet send from \textbf{node 1} to \textbf{node n} over \textbf{nodes 2 to n-1} cannot be received at a faster rate than the lowest bandwidth on the path of the packet (bottleneck) [Figure \ref{fig:transmission delay}].\[Rate_{max} \leftarrow \min\{bandwidth_{i}\mid i\in\{1,...,n-1\}\}\]

\item
Each link creates an intentional gap between two packets being sent. This pause is often necessary to allow the receiver to prepare for a new package, however, depending on the transmission protocol, other purposes may apply. This kind of delay is known as \textit{Inter-Packet delay}.

@@

It is important to understand that delay has a direct impact on congestion. When delay increases, the networks throughput can decrease as packets will stay inside the network for a longer time. If the network however continues to receive new packets at constant rate, more and more packets will have to be stored, resulting in congestion. Generally speaking, to tackle congestion, firstly delay can be reduced, and secondly the networks storages, the buffers, can be improved regarding efficient queue management and buffer size. Finally, transmission protocols can also react to congestion and adapt their send rate to recover from congestion [\nameref{sec:ca}] and even control congestion [\nameref{sec:cc}].
\\Whereas processing delay is almost not noticeable and transmission and inter-packet delay could be improved by upgrading cables or transmission protocols, queuing delay has various causes and therefore is the  most difficult one to reduce. There are many different ways to implement buffers in a network node and the effectiveness of each implementation varies from where it is used. Traffic in residential network nodes for example differs from nodes used in large data-centers in every aspect, thus hardware and software in those components should be adjusted accordingly.
\\Sizing buffers is a complex task. On the one hand read and write times should be as short as possible to minimize additional delay. On the other hand buffers should not be too small, as networks need to be robust for future changes and therefore nodes should store lots of data if needed, while also keeping in mind the requirements of congestion detection algorithms as described in section \nameref{sec:ca} in order to adjust for congestion appropriately \cite{staff2012bufferbloat}. Unifying those goals is very challenging as it leads to increasing costs and physical space occupied by those buffers. Additionally, a 2004 paper by Appenzeller et al. \cite{appenzeller2004sizing} found that 99\% of buffers in network nodes, which were sized on a rule-of-thumb rule based on the TCP protocol \cite{villamizar1994high}, could be removed without losing, but rather gaining performance. Clearly, balancing speed and size is difficult and thus buffer sizes should be adjusted to their area of use accordingly.
\\This isn't the only major cause of queuing delay. Without proper buffer management even nodes with well-sized buffers will experience performance loss. Proper queue management is able to compensate for the unpredictable fluidity of network traffic. It is a difficult task to buffer the right amount of packages to keep transmission flow consistent over time.
As the number of packets being transmitted grows, the \textit{throughput} of the network node increases until packets are being transmitted at bottleneck rate. After this, the transmission rate cannot increase anymore, which leads to packets being buffered along their path to prevent them from being lost \cite{gettys2012bufferbloat}. If the incoming amount of traffic keeps to exceed the amount of outgoing traffic, those buffers will fill up (\textit{over-buffering}), which will eventually lead to \textit{bufferbloat} \cite{Allman13commentson,6886125}.

\subsection*{Bufferbloat}
Bufferbloat refers to frequently full and excessively large buffers in internet nodes. This leads to partly unnecessary network delay and damages congestion avoidance algorithms, such as those found in classic TCP, which will be explained in the section \nameref{sec:ca} \cite{gettys2012bufferbloat,chen2014bufferbloat}. Buffering is needed to provide space to queued packets waiting for transmission resulting in a minimization of packet loss. In the past however, buffer sizes were smaller than today's thus less packets could be queued and more were dropped, signaling transmission protocols the presence of congestion in this node early in. With memory getting cheaper over the years, buffers saw an increase in size. This has the negative impact, that more packets can be queued before any are dropped, thus those adjustments of congestion avoidance protocols kick in later than usual. Therefore, buffers are being filled in the whole network which results in an extreme increase in network delay (\textit{latency}) and fluctuation (\textit{jitter}) \cite{gettys2012bufferbloat,staff2012bufferbloat,chen2014bufferbloat}.
\\So Services, which require low latency at a high data-rate, such as streaming, will experience a decrease in quality.

\section{Congestion Avoidance}
\label{sec:ca}
As already mentioned, congestion detection algorithms make up an important part in congestion avoidance protocols. \textit{TCP} is such a protocol, based on the transport layer, which adapts its send rate according to available network bandwidth \cite{1209197,jacobson1992tcp}. To fully understand, how TCP's congestion avoidance algorithm operates, one needs to understand how TCP works. In the following TCP's congestion avoidance algorithm will be explained using the TCP Specification \cite{ietf-tcpm-rfc793bis-16,postel1981transmission}, also being refered to as \textit{classic TCP}. Note, there are also other TCP implemtations improving upon it. The underlying concept however is the same.
\\TCP lives on the transport layer. It communicates with the application layer via sockets, splits up data into packets (and vice versa), and transmits those packages over the network with the help of the internet protocol, thus TCP is often also referred to as \textit{TCP/IP}. It is the interface between network components and computer programs. Each of those packets have different purposes, but underlay the same structure - the protocol header. The TCP header is split up into multiple segments [Figure \ref{fig:TCP_Header}]:
\begin{figure}
\centering
\begin{bytefield}[bitheight=2.2\baselineskip ,bitwidth=0.635\baselineskip]{32}
	\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
	\bitbox{32}{Sequence Number} \\
	\bitbox{32}{Acknowledgment Number} \\
	\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} &
\bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} &
\bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} &
\bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} &
\bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} &
\bitbox{16}{Window Size} \\
\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
\bitbox{32}{Options} \\
\bitbox{32}{Data}
\end{bytefield}
\caption{Basic structure of the TCP header. It is split up into multiple 32-bit blocks. The data block contains all the application data to be transmitted \cite{jacobson1992tcp,ietf-tcpm-rfc793bis-16,huston2000tcp}.}
\label{fig:TCP_Header}
\end{figure}
\begin{itemize}
\item \textit{Source Port} and \textit{Destination Port} specify the ports on which TCP should communicate with the application layer.

@@

\subsection{Perceiving Congestion}
\label{sec:P_C_marker}
Both partners of a connection are both sender and receiver, as both transmit packets. In the following however, we will simplify this and specify one partner as the sender, the other as the receiver. The sender keeps track of the time it takes from when a package sent until the acknowledgment is received. This is also known as \textit{round-trip time} (RTT). Comparing the current RTT to previously measured RTTs can inform the sender about the current state of congestion. Some variances of TCP make use of this feature in their congestion avoidance algorithms, yet classic TCP does not \cite{huston2000tcp,jacobson1995congestion}. RTT also plays a very important role in the detection of packet loss, which further implies congestion. The sender side keeps track of the RTT while it is being measured and compares it to the \textit{retransmission timeout} (RTO) \cite{jacobson1992tcp}. RTO time is being calculated by estimating the mean and variance of the RTT. Whenever packets are not acknowledged within this RTO interval, the TCP sender regards this packet as lost and will retransmit its data. Therefore, it is essential to TCP performance, that the RTO is being determined accurately, as a too low RTO could prompt unnecessary retransmissions and too high RTO could slow down the application TCP is serving \cite{huston2000tcp,jacobson1995congestion}. To implicitly notify TCP that congestion occurs, congested network nodes will occasionally drop packets after a certain threshold of queued packets has been exceeded. Usually, this is quite smaller than the maximum queue length, therefore avoiding actions can be taken in time \cite{ramakrishnan1999proposal}.

@@

\subsection{Adjusting for Congestion}
\label{sec:AfC}
To prevent congestion altogether, packet flow inside a network would have to be conservative. A connection in this state is also being described as being \textit{in equilibrium} \cite{jacobson1992tcp}. This simply means, the networks works at full load whilst no additional delay or congestion is being created. However, it is quite impossible to keep a full network in this state forever. The goal of TCP's congestion avoidance algorithms, which are based on an \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle, is to react to the current state of congestion and achieve equilibrium. Those algorithms can be categorized into three sub-algorithms, that all have the goal to keep the connection in equilibrium, no matter how congestion was perceived. Here, a variable called \textit{congestion window} (cwnd) has an important role. It describes the maximum amount of data sent during an average RTT. An additional mechanism of keeping a connection in equilibrium is, that each acknowledgment packet received by the sender opens the space for a new packet to be sent into the network \cite{jacobson1992tcp}. The principles TCP uses to adjust for congestion are based on \textit{Additive-Increase Multiplicative-Decrease} (AIMD) principle.

\section{Network Assisted Congestion Control}
\label{sec:cc}
The problem that comes with classic TCP is that the congestion detection algorithms are very implicit. Packet loss for example may have different causes other than network congestion, such as simple cable interference, however, the congestion avoidance algorithm will still interpret it as congestion, thus possibly resulting in unnecessary network delay. Therefore, exactly determining whether the network is congested or not is crucial to keep a network stable. Here, the network layer comes into play. Because most communication should be hidden along the path of the packet, all network nodes will only communicate on the network layer (where IP lives), but not above (where TCP lives) \cite{ramakrishnan1999proposal}.\\
In 2001 a new method was standardized to explicitly notify TCP on congestion. \textit{Explicit Congestion Notification} (ECN) is an extension to TCP and IP, allowing end-to-end congestion notification, while also minimizing the amount of packets dropped. This is done by modifying the IP header to be able to carry one more bit - the \textit{Congestion Experienced} flag (CE) \cite{ramakrishnan1999proposal}. Instead of dropping packets occasionally after the amount of queued packets exceed a threshold [\nameref{sec:P_C_marker}], all packets supporting ECN will now be marked (setting the CE bit in the IP header) [Figure \ref{fig:ecn}]. Thus, congestion can be described explicitly resulting in better short- and long-term network performance. Finally, if a data packet has been marked, the according acknowledgment packet will also be marked to inform the sender of congestion \cite{ramakrishnan1999proposal,ramakrishnan2001addition,10.1145/205511.205512}.\\
To sum up so far, without ECN, classic TCP would just drive the network into congestion, and then recover from it. The network generally speaking will at least experience some jitter (fluctuating network delay). With ECN however, packets in congested nodes will be marked, explicitly informing the TCP sender that the network is congested. As a result, avoiding actions can be taken immediately. In addition, the need to retransmit packets would decrease drastically, as no packets are being dropped anymore to indicate congestion. Therefore, this central algorithm is also being called \textit{Congestion Control} \cite{ramakrishnan1999proposal,ramakrishnan2001addition}.

@@

\subsubsection*{XCP Sender}
As with TCP, the XCP sender maintains a congestion window, \textit{cwnd}, and an estimate of the round trip time, \textit{RTT}. On packet departure, the \textit{H\_cwnd} and \textit{H\_rtt} fields will be initialized with \textit{cwnd} and \textit{RTT} accordingly. The initialization value of the \textit{H\_feedback} field represents a desired send rate $r$:
\[H\_feedback \leftarrow \frac{r \cdot RTT - cwnd}{cwnd}\]
This is an improvement to TCP's adjustment algorithms as the desired rate can be reached after one RTT.\\
The other task an XCP sender has is to adjust the congestion window according to the networks feedback. When $s$ is the packet size, the congestion window variable would be adjusted by following formula:
\[cwnd \leftarrow \max(cwnd + H_feedback, s)\]
In addition to responding to direct feedback, the XCP sender still needs to handle packet loss, though this is very similar to what TCP does \cite{katabi2002congestion}.

@@

\subsubsection*{XCP Router}
The main task of network nodes in XCP, or XCP routers, is to inform the XCP sender of the current state of congestion. This is done by computing feedback at every node. The objective of XCP is to prevent the queues building up to a point at which packets have to be dropped \cite{katabi2002congestion}.\\
Feedback is computed by two controllers. The \textit{Efficiency Controller} (EC) tries to maximize link utilization while minimizing queue lengths and packet drop rates. When the XCP routers link is underused, positive feedback should be sent. If the link is congested however, negative feedback should be sent in order to keep queue lengths small resulting in no packet drops \cite{katabi2002congestion,1498331}.\\
To achieve efficiency and equilibrium, this feedback will be allocated to single packets as \textit{H\_feedback}. However, the EC does not decide which packets should carry the feedback. Hence, the \textit{Fairness Controller} (FC) comes into play. As TCP, FC relies on the AIMD principle [\nameref{sec:AfC}]. If the EC calculated positive feedback, it will be split up between all packets equally. If the feedback is negative however, it will be divided between packets in proportion to their current send-rates, which are determined by \textit{H\_cwnd} and \textit{H\_rtt} \cite{katabi2002congestion,1498331}.\\
Decoupling the feedback computation allows XCP to quickly acquire the maximum send rate and achieve equilibrium, all while the network allocates bandwidth to all XCP senders equally \cite{katabi2002congestion}.

\subsection{Evaluation of Network Assisted Congestion Control}
As you can see, including the network layer to detect congestion can have great performance improvements. 'Understanding XCP: Equilibrium and Fairness' has shown through simulations that many network topologies could benefit from XCP as it clears queues in equilibrium, thus resulting in improved network performance under load \cite{1498331}. Another paper has shown that XCP outperforms ordinary TCP in nearly any environment related to average queue lengths, efficiency, and throughput \cite{katabi2002congestion}.
\\However, all this performance comes at a cost. Since network nodes have to actively detect congestion, ECN, XCP, and other network assisted congestion control protocols rely on active queue management in those nodes. Although there is some hardware being produced today that meet ECNs requirements, a large part of internet hardware isn't capable of doing so. As many routers do not run on software but hardware only to improve performance, they can only handle specific packets. Whenever they cannot recognize a packet, it will simply be dropped \cite{katabi2002congestion}. So, although ECN packets differ from ordinary TCP/IP packets by only one bit, not all routers can handle this change and will drop those packets. This is the reason you do not see those protocols in wide area networks (WAN) too often, let alone the internet.
\\Nevertheless, whenever the implementation of XCP, ECN, or similar protocols is not a challenge, you can implement network assisted congestion control with ease and see great performance improvements. Especially smaller networks with high traffic could definitely benefit from those technologies [\nameref{sec:DCTCP}].

@@

\subsection{DCTCP}
\label{sec:DCTCP}
A lot of internet services run in data centers with large amounts of computers communicating with each other, resulting in a lot of traffic inside the data center. Optimizing those data centers networking capabilities can result in an increase in application productivity. In 2010 a paper proposed \textit{DCTCP}, a variation of TCP using ECN to create a data transportation protocol, which especially takes hardware purposed for use in those environments into account \cite{10.1145/1851275.1851192}.\\
The general problem that appears in high-traffic environments is the so-called \textit{TCP Incast} problem. It describes a drastic reduction in application throughput. To understand where TCP incast originates, you fist need to understand how most data centers are built. A typical data center contains a set of racks, each holding a large number of servers, the switches, that connect those racks, and connecting links, that connect those switches to other parts of the data center or the internet [Figure \ref{fig:dc}]. This design pattern is also known as \textit{Partition/Aggregate} and is most commonly used in data centers. It however results in a high 'fan-in', as multiple servers (leaves) are connected to just one link (root) \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}. In addition, most applications require high bandwidth, low latency connections, also enabling them to send many parallel requests to other servers. Combining those requirements with the design pattern, it becomes clear, that those switches are being challenged. Besides, many data center switches have relatively small, shared buffers. Therefore, it is common that queue lengths stay high for a prolonged period of time, resulting in a collapse of throughput \cite{10.1145/1851275.1851192,10.1145/1592681.1592693}.\\
The goal of DCTCP is to keep data center traffic constantly flowing, which means meeting the requirements applications take while keeping relatively small buffer sizes inside switches in mind. To achieve this, DCTCP implements ECN. Unlike TCP however, if a DCTCP sender registers congestion, the window size will be reduced by a fraction of the amount of marked packets it received. Therefore, excessive link underutilization can be prevented. In addition, the DCTCP receiver will only send one ACK packet for every $m$ consecutive data packet, where all $1,...,m$ packets before had the same CE-flag configuration. Here, $m$ is a variable initialized while the DCTCP connection is being established and cannot be changed without renegotiation. Therefore, fewer packets are sent through the network, but the TCP sender is still able to identify lost packets and register congestion \cite{10.1145/1851275.1851192}.
\\Being bandwidth greedy transmission protocol, data centers using DCTCP can see a performance drop in nodes that use other transmission protocols such as TCP or ECN. As DCTCP tries to utilize the links as much as possible, non-greedy protocols will be neglected and will see a drop in performance. Thus, data centers should avoid using DCTCP in addition to other protocols to fully make use of DCTCP's capabilities.
\\To sum up, being able to react to congestion immediately enables DCTCP to reduce buffer overflows and timeouts, thus largely resolving incast problems. The paper proposing DCTCP has found, that their protocol increases traffic by 110\% while using 90\% less buffer space compared to TCP \cite{10.1145/1851275.1851192}.
